---
title: "BDA - Assignment 1"
output: 
  pdf_document
---

```{r setup, include=FALSE}
# This chunk sets echo = TRUE as default, that is print all code.
# knitr::opts_chunk$set can be used to set other notebook generation options, too.
# include=FALSE inside curly brackets makes this block not be included in the pdf.
knitr::opts_chunk$set(echo = TRUE)
```

# Loaded packages

```{r}
# To install aaltobda, see the General information in the assignment.
library(aaltobda)
```


# Basic probability theory notation and terms

probability: the chance that a given event will occur (or is true), quantified between 0 and 1 where 0 is impossibility and 1 is certainty.

probability mass: probability distributions of discrete random variables

probability density: probability distributions of continuous random variables

probability mass function (pmf): function that provides the probability that a discrete random variable is exactly equal to a given value.

probability density function (pdf): function that provides the probability that a continuous random variable is exactly equal to a given value.

probability distribution

discrete probability distribution

continuous probability distribution

cumulative distribution function (cdf)

likelihood

# Basic computer skills

a) Plot the density function of Beta-distribution, with mean $\mu$ = 0.2 and variance $\sigma^2$ = 0.01.

```{r}
x <- seq(0,1, length = 21)
mu <- 0.2
sig <- 0.01

alpha <- mu*((mu*(1-mu)/sig)-1)
beta <- (alpha*(1-mu))/mu
beta_dist <- dbeta(x, shape1 = alpha, shape2 = beta)
plot(beta_dist, type = "l", ylab = "")
```

b) Take a sample of 1000 random numbers from the above distribution and plot a histogram of the results.

```{r}
beta_vals <- rbeta(1000, alpha, beta)
hist(beta_vals)
```

c) Compute the sample mean and variance from the drawn sample. Verify that
they match (roughly) to the true mean and variance of the distribution.

```{r}
# sample mean and variance from the drawn sample
mean(beta_vals)
mu

var(beta_vals) 
sig
# they match!
```

d) Estimate the central 95% probability interval of the distribution from the
drawn samples.

```{r}
quantile(beta_vals, probs = c(0.025, 0.975))
```

# 3. Bayes' theorem

$$ P(A|B) = \frac{P(B|A)P(A)}{P(B)} $$

where 

$ P(A|B) $ is the conditional probability that A will occur given that B is true
$ P(B|A) $ is the conditional probability that B will occur given that A is true
$ P(A) $ and $ P(B) $ is the marginal probability of observing A and B respectively

Here,
$A$ = $Cancer$ = Subject has lung cancer
$B$ = $Positive$ = Test gives positive

So, we are looking for:
$$ P(Cancer | Positive) = \frac{P(Positive | Cancer)P(Cancer)}{P(Positive)}$$

$$ P(Cancer | Positive) = \frac{P(Positive | Cancer)P(Cancer)}{P(Positive|Cancer)P(Cancer)+P(Positive|NoCancer)P(NoCancer)}$$

We know that:

$P(Positive | Cancer) = 0.98$

$P(Cancer) = 1/1000 = 0.001$

$P(Positive|NoCancer) = 1 - 0.96 = 0.04$

$P(NoCancer) = 999/1000 = 0.999 $

so $P(Positive) = 0.98*0.001 + 0.04*0.999$

Meaning,

$$ P(Cancer | Positive) = \frac{0.98 * 0.001}{0.04}$$
```{r}
((0.98*0.001)/(0.98*0.001 + 0.04*0.999))*100
```

Which is very low. This means the joint probability of someone getting a positive test and having lung cancer is ~4%, which should be improved before getting the test to market.

# 4. 

### a)

```{r}
boxes <- matrix(c(2,4,1,5,1,3), ncol = 2, dimnames = list(c("A", "B", "C"), c("red", "white")))

p_red <- function(boxes){
  
  pA <- 0.4
  pB <- 0.1
  pC <- 0.5

  prA <- boxes["A", "red"]/sum(boxes["A",])
  prB <- boxes["B", "red"]/sum(boxes["B",])
  prC <- boxes["C", "red"]/sum(boxes["C",])

  pr <- pA*prA + pB*prB + pC*prC

  return(pr)

}
```

### b)

```{r}
p_box <- function(boxes){

  pA <- 0.4
  pB <- 0.1
  pC <- 0.5

  prA <- boxes["A", "red"]/sum(boxes["A",])
  prB <- boxes["B", "red"]/sum(boxes["B",])
  prC <- boxes["C", "red"]/sum(boxes["C",])

  pr <- pA*prA + pB*prB + pC*prC

  pr_A <- pA*prA/pr
  pr_B <- pB*prB/pr
  pr_C <- pC*prC/pr

  return(c(pr_A, pr_B, pr_C)) 

}

```

# 5.

```{r}
p_identical_twin <- function(fraternal_prob, identical_prob){
  identical_prob/(fraternal_prob/2 + identical_prob)
} 
p_identical_twin(identical_prob = 1/400, fraternal_prob = 1/150)
```

